\documentclass[11pt]{article}

% Edit this customize for an instructor

\newcommand{\instructorpronoun}[1]{his}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

% Define the headers and footers

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 100}
\lhead{Syllabus}
\rfoot{Page \thepage}
\lfoot{Fall 2018}

% Use elastic spacing around the headers

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\syllabustitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 100\\Computational Expression\\
      Fall 2018\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\syllabustitle{Syllabus}

\vspace*{-.1in}
\subsection*{Course Instructor}
Dr.\ Gregory M.\ Kapfhammer\\
\noindent Office Location: Alden Hall 108 \\
\noindent Office Phone: +1 814--332--2880 \\
\noindent Email: \url{gkapfham@allegheny.edu} \\
\noindent Twitter: \url{@GregKapfhammer} \\
\noindent Web Site: \url{https://www.gregorykapfhammer.com/}

\subsection*{Instructor's Office Hours}

\vspace*{-.05in}

\begin{itemize}
  \itemsep0em

  \item Monday: 10:00 am--11:00 am and 2:30 pm--3:30 pm (15 minute time slots)

  \item Tuesday: 4:30 pm--5:00 pm (15 minute time slots)

  \item Wednesday: 10:00 am--11:00 am and 2:30 pm--3:30 pm (15 minute time slots)

  \item Thursday: 11:00 am--12:00 noon and 4:30 pm--5:00 pm (15 minute time slots)

  \item Friday: 10:00 am--11:00 am (15 minute time slots)

\end{itemize}

\noindent To schedule a meeting with me during my office hours, please visit my
web site and click the ``Schedule'' link in the top left-hand corner. Now, you
can browse my office hours or schedule an appointment by clicking the correct
link and then reserving an open time slot. Students are also encouraged to post
appropriate questions to a channel in Slack, which is available at
\url{https://CMPSC100Fall2018.slack.com/}, and monitored by the instructor and
the teaching assistants.

\subsection*{Course Meeting Schedule}

Lecture, Discussion, and Group Work Session: Monday and Wednesday 1:30 pm--2:20 pm \\
Practical Session: Friday 1:30 pm--2:20 pm \\
Laboratory Session: Tuesday 2:30 pm--4:20 pm \\
Final Examination: Friday, December 14, 2018 at 7:00 pm

\subsection*{Course Description}

\begin{quote}

An introduction to the principles of computer science with an emphasis on
creative expression through the medium of a programming language. Participating
in hands-on activities that often require teamwork, students learn the
computational structures needed to solve problems and produce artifacts such as
interactive games and computer-mediated art and music. Students also learn how
to organize and document a program's source code so that it effectively
communicates with the intended users and maintainers. Additionally, the
introduction includes an overview of the discipline of computer science and
computational thinking. During a weekly laboratory session students use
state-of-the-art technology to complete projects, reporting on their results
through both written documents and oral presentations.
Prerequisite: Knowledge of elementary algebra.
Distribution Requirements: ME, SP.\@ \\

\end{quote}

\subsection*{Course Objectives}

The process of implementing and evaluating correct and efficient software
involves the application of many interesting theories, techniques, and tools. In
addition to learning problem solving and computational thinking skills, this
class will teach students how to use, design, implement, and test algorithms in
the object-oriented programming language called Java. Students will also learn
about fundamental concepts such as data types, conditional logic, and iteration
while also discovering how to use single-dimension, multi-dimensional, and
extendable arrays and to implement graphical applications. Students will also
gain hands-on experience in the use, design, implementation, and testing of
software during the laboratory and practical sessions and a final project. Along
with learning more about how to effectively work in a team of diverse software
developers, students will enhance their ability to write and present ideas about
software in a clear, concise, and compelling fashion. Students will also develop
an understanding of the fascinating connections between computer science and
other disciplines in the social and natural sciences and the humanities.

\subsection*{Performance Objectives}

At the completion of this semester, students should have a strong grasp of the
basics of the object-oriented programming paradigm and an ever-deepening
knowledge of topics like conditional logic, iteration, recursion, exceptions,
and graphics programming. Also, students should be able to handle many of the
important, yet accidental, aspects of implementing programs in the Java
language. For instance, students should be comfortable with the use of Vim (and
other text editors such as Atom) and understand how to use software tools such a
GitHub, Gradle, and GatorGrader. Students should have a toolkit of programming
language constructs that they can use to respond to the challenges that they
encounter during the development and evaluation of software. Finally, students
should demonstrate the ability to use both in-person discussions and
cutting-edge software tools to effectively communicate and collaborate with a
group of diverse team members.

\vspace*{-.05in}

\subsection*{Required Textbook}

\noindent{\em Java Software Solutions: Foundations of Program Design}. John
Lewis and William Loftus, Ninth Edition, ISBN:\ 0134462025, ISBN--13:\
978-0134462028, 800 pages, 2017. \\ (References to the textbook are abbreviated
as ``JSS'' on the course web site and the syllabus).

\vspace*{.1in}

\noindent Students who want more information about Java or project ideas may
consult this free book:

\noindent{\em Think Java: How to Think Like a Computer Scientist}. Allen Downey
and Chris Mayfield. First Edition, ISBN--10: 1491929561, ISBN--13:
978-1491929568, 252 pages, 2016.

\vspace*{.1in}

\noindent Students who want to improve their technical writing skills may
consult this suggested book:

\noindent{\em Writing for Computer Science}. Justin Zobel. Second Edition,
ISBN--10: 1852338024, ISBN--13: 978--1852338022, 270 pages, 2004.

\vspace*{.1in}

\noindent Along with reading the required textbook, you may be invited to study
additional articles from a wide variety of conference proceedings, scientific
journals, and the popular press.

\vspace*{-.05in}

\subsection*{Course Policies}

\subsubsection*{Grading}

The grade that a student receives in this class will be based on the following categories. All of these percentages are
approximate and, if the need to do so presents itself, it is possible for the course instructor to change the assigned
percentages during the academic semester.

\begin{center}
  \begin{tabular}{c c}
  \begin{tabular}{ll}
    Class Participation        & 5\%  \\
    SEED Project Participation & 5\%  \\
    Midterm Examination        & 15\% \\
    Final Examination          & 15\% \\
  \end{tabular} &
  \begin{tabular}{ll}
    Mastery Quizzes            & 10\% \\
    Laboratory Assignments     & 30\% \\
    Practical Assignments      & 10\% \\
    Final Project              & 10\%
  \end{tabular}
  \end{tabular}
\end{center}

\vspace*{-.05in}

\noindent
These grading categories have the following definitions:

\vspace*{-.05in}

\begin{itemize}

  \item {\em Class Participation\/}: All students are required to actively
    participate during all of the course sessions. Your participation will take
    forms such as answering questions about the reading assignments, asking
    constructive questions of group members, giving presentations, and leading a
    discussion. You also must regularly participate in the discussions in the
    course's Slack workspace. A student may request feedback on and will receive a
    final grade for this category.

  \item {\em SEED Project Participation\/}: As part of an ongoing project that
    publishes educational discussions with software engineers, I will regularly
    publish, on my blog available at
    \url{https://www.gregorykapfhammer.com/blog/}, an interview with a software
    developer in industry. At the start of an in-class practical session on a
    Friday, every student should read the blog post and tweet a response to
    that week's interview. Along with linking to the web site of the blog post
    and mentioning \url{@GregKapfhammer}, your tweet should share what you
    learned from the interviewee. Interested students should also connect with
    the interviewee through LinkedIn.

  \item {\em Mastery Quizzes\/}: The two quizzes will cover all of the material
    in their associated module(s), as outlined on the review sheet. While the
    second quiz is not cumulative, it will assume a basic understanding of the
    material that was the focus of the first quiz. The date for each of the
    quizzes will be announced at least one week in advance of the scheduled
    date. Unless prior arrangements are made with the instructor, all students
    will be expected to take these two quizzes on their scheduled date and to
    complete the quizzes in the stated period of time.

  \item {\em Midterm Examination\/}: The midterm is an hour-long cumulative test
    covering all of the material from the class, practical, and laboratory
    sessions, as outlined on the review sheet. Unless prior arrangements are
    made with the course instructor, all students will be expected to take this
    test on the scheduled date and complete the test in the stated period of
    time.

  \item {\em Final Examination\/}: The final examination is a three-hour
    cumulative test, as outlined on the review sheet. Unless there are
    documented severe and extenuating circumstances, students must take this
    examination at the date and time stated on the first page of the syllabus.

  \item {\em Laboratory Assignments\/}: These assignments invite students to
    explore different techniques for designing, implementing, evaluating, and
    documenting software solutions to challenging problems that often have a
    connection to real-world concerns. Some of the assignments will require
    students to conduct experiments and collect, analyze, and write about data
    sets. To best ensure that students are ready to develop software in both
    other classes at Allegheny and after graduation, students will complete
    assignments both on an individual basis and in teams. When teamwork is
    required, the instructor will often assign individuals to teams.

  \item {\em Practical Assignments\/}: Graded on a credit/no-credit basis, these
    assignments allow students to practice the technical skills that they
    learned in previous class and laboratory sessions.

  \item {\em Final Project\/}: This project will furnish you with the
    description of a challenging problem and ask you to design, implement,
    describe, and orally present a correct and carefully evaluated solution.
    Completion of the final project will require you to apply all of the
    knowledge and skills that you have acquired during the semester to solve a
    technical problem and, whenever possible, make your solution and results
    publicly available in a free and open fashion.

\end{itemize}

\subsubsection*{Assignment Submission and Evaluation}

All assignments will have a stated due date. Electronic versions of the
laboratory, practical, and final project assignments must be submitted to a
student's GitHub repository; students will learn how to use version control with
GitHub during the first laboratory and practical sessions. No credit will be
awarded for any course work that is not submitted to your GitHub repository with
the required name. Unless specified otherwise, all laboratory assignments must
be turned in at the beginning of the session that is one week after the day the
assignment was released. Practical assignments will be due at the start of the
next class session unless otherwise stated. Without making special arrangements
with the course instructor, no work will be accepted after the published
deadline. Using a report that the instructor shares with you through the commit
log in GitHub, you will privately received a grade for and feedback on each
assignment. Your grade will be a function of whether or you not completed
correct work and submitted it by the deadline. Other factors (e.g., the quality
of your source code and technical writing) will also influence your assignment
grade.

\vspace*{-.1in}

\subsubsection*{Course Attendance}

It is mandatory for all students to attend all of the class, practical, and
laboratory sessions. If, due to extenuating circumstances, you will not be able
to attend a session, then, whenever possible, please see the course instructor
at least one week in advance to describe your situation. Students who miss more
than five unexcused sessions will have their final grade in the course reduced
by one letter grade. Students who miss more than ten of the aforementioned
events will fail the course.

\vspace*{-.05in}

\subsubsection*{Use of Laboratory Facilities}

Throughout the semester, we will employ many different software tools that
computer scientists use during the design, implementation, and evaluation of
computer software. Since it is unlikely that a student's unmodified computer
will have the software tools that correctly support the completion of the
laboratory assignments, unless there are documented extenuating circumstances,
students are advised to complete all of their work for this course while using
the laboratory facilities.

\vspace*{-.1in}

\subsubsection*{Class Preparation}

In order to minimize confusion and maximize learning, students must invest time
to prepare for the class discussions, lectures, and practical sessions. During
the class periods, the course instructor will often pose challenging questions
that could require group discussion, the creation of a program or data set, a
vote on a thought-provoking issue, or a group presentation. Only students who
have prepared for class by reading the assigned material and reviewing the
current laboratory and practical assignments will be able to effectively
participate in these discussions.

More importantly, only prepared students will be able to acquire the knowledge
and skills that they need to be successful in this course, subsequent courses,
and the field of computer science. In order to help students remain organized
and to effectively prepare for classes, the course instructor will maintain a
class schedule with reading assignments and presentation slides. During the
class sessions students will also be required to download, use, and modify
programs and data sets that are made available through means such as the course
web site and a GitHub repository.

\subsubsection*{Seeking Assistance}

Students who are struggling to understand the knowledge and skills developed in
a class, laboratory, or practical session are encouraged to seek assistance from
the course instructor, the teaching assistants, or the departmental tutors.
Throughout the semester, students should, within the bounds of the Honor Code,
ask and answer questions on the Slack workspace for our course; please request
assistance from the instructor, teaching assistants, and tutors first through
Slack before sending an email. Students who need the course instructor's
assistance must schedule a meeting through \instructorpronoun{} web site and
come to the meeting with all of the details needed to discuss their question.

\subsubsection*{Using Email}

Although we will primarily use Slack for class communication, I will sometimes
use email to send announcements about important matters such as changes in the
schedule. It is your responsibility to check your email at least once a day and
to ensure that you can reliably send and receive emails. This class policy is
based on the statement about the use of email that appears in {\em The Compass},
the College's student handbook; please see the instructor if you do not have
this handbook.

\subsubsection*{Honor Code}

The Academic Honor Program that governs the entire academic program at Allegheny College is described in the Allegheny
Academic Bulletin. The Honor Program applies to all work that is submitted for academic credit or to meet non-credit
requirements for graduation at Allegheny College. This includes all work assigned for this class (e.g., examinations,
laboratory assignments, and the final project). All students who have enrolled in the College will work under the Honor
Program. Each student who has matriculated at the College has acknowledged the following pledge:

\vspace*{-.11in}
\begin{quote}
  I hereby recognize and pledge to fulfill my responsibilities, as defined in the Honor Code, and to maintain the
  integrity of both myself and the College community as a whole.
\end{quote}
\vspace*{-.11in}

\noindent It is understood that an important part of the learning process in any course, and particularly one in
computer science, derives from thoughtful discussions with teachers and fellow students. Such dialogue is encouraged.
However, it is necessary to distinguish carefully between the student who discusses the principles underlying a problem
with others and the student who produces assignments that are identical to, or merely variations on, someone else's
work. While it is acceptable for students in this class to discuss their programs, data sets, and reports with their
classmates, deliverables that are nearly identical to the work of others will be taken as evidence of violating the
\mbox{Honor Code}.

\subsubsection*{Disability Services}

The Americans with Disabilities Act (ADA) is a federal anti-discrimination statute that provides comprehensive civil
rights protection for persons with disabilities. Among other things, this legislation requires all students with
disabilities be guaranteed a learning environment that provides for reasonable accommodation of their disabilities.
Students with disabilities who believe they may need accommodations in this class are encouraged to contact Disability
Services at 332--2898. Disability Services is part of the Learning Commons and is located in Pelletier Library.
Please do this as soon as possible to ensure that approved accommodations are implemented in a timely fashion.

\subsection*{Welcome to an Adventure in Computer Science}

In reference to software, Frederick P.\ Brooks, Jr.\ wrote in chapter one of {\em The Mythical Man Month}, ``The magic
of myth and legend has come true in our time.'' Software is a pervasive aspect of our society that changes how we think
and act. Efficient and correct software also has the potential to positively influence the lives of many people.
Moreover, the design, implementation, evaluation, and documentation of software are exciting and rewarding activities!
At the start of this class, I invite you to pursue, with great enthusiasm and vigor, this adventure in computer science.

\end{document}
